---
title: "HomeWork 2"
author: "ivan Ojiambo"
date: "September 24, 2017"
output:
  html_document:
    df_print: paged
---

#Exe1-2
The data was seperated by tabs and it has 12 columns and 32561 rows.  I have replaced all the missing values with NA
```{r}

unclean_data <- read.csv( "data.csv",  sep = "\t", strip.white = TRUE, na.strings = c("NA", "?"));
#type of data
str(unclean_data)

```
I have removed Column X because  it does not make sense since the rows are always numbered automatically
```{r}
unclean_data[,c("X")] <- NULL

```
## 1. Analysing column Age
There are  97 people without age value.
```{r}

sum(is.na(unclean_data$age))
```

The minimum  and maximum age in the data is -57 years and  320 years respectively, I have subsituted these values  with  missing value because its unrealistic to have negative vale as age neither is it realistic to be over 320 years. 
i have assumed one can not be over 100 years and below 1 year

```{r}
summary(unclean_data$age)

unclean_data$age[unclean_data$age < 1] <- NA
unclean_data$age[unclean_data$age  >100] <- NA




```


```{r}
hist(unclean_data$age, xlab = "Age", ylab = "Frequency", main = "Age Distribution")
```
 Age is a ratio

## 2. Analysing Workclass Column
There are 1836 people who are not working.
```{r}
sum(is.na(unclean_data$workclass))

```

There were 10 people who had **privat** as there working class, I decided to add them to **Private** work class. I assumed it was  typo since privat was not making sense.

```{r}
library(plyr)
 revalue(unclean_data$workclass, c("privat" = "Private")) -> unclean_data$workclass

counts <- table(unclean_data$workclass)
barplot(counts,  main = "Work class Distribution", xlab = "Work class", ylab = "freqency", cex.names = 0.6 ) 

``` 

Work class is of nominal data type.


## 3. Analysing Column Education
Everyone atlest went to school. ie there is no missing value
```{r}
library(plyr)
sum(is.na(unclean_data$education))
counts <-table( unclean_data$education)
 
unclean_data$education[unclean_data$education =="12th"] <- "1st-12th"

barplot(counts, main = "Education Distribution", xlab = "Education Level", ylab = "frequency", cex.names = 0.6)

```
Education is of ordinal type


## Analysing Attribute occupation
There are 1843 people who do not have occupation
```{r}
sum(is.na(unclean_data$occupation))
```


```{r}
counts <- table(unclean_data$occupation)

barplot(counts,  main = "Occupation Distribution by population", xlab = "Jobs", cex.names = 0.6 ,ylab = "frequency")
```

occupation is of nominal data type.


## 5. Analysing capital gain

There are no missing value for capital gain. The maximum value is 99999 and the minimum value is 0

```{r}
summary(unclean_data$capital.gain)

plot(density(unclean_data$capital.loss), main = "Kernal density plots showing values of capital Loss")

```

Capital gain attribute is of ratio data type.

## 6. Analysing capital loss

There is no missing value for capital loss and the maximum value is 4356 and the minimum value is 0
```{r}
library(plyr, warn.conflicts = FALSE)
summary(unclean_data$capital.loss)

plot(density(unclean_data$capital.gain), main = "Kernal density plots showing values of capital gain")
```

Capital loss attribute is of ratio data type.


## 7. Analysing column Native country
There are 583 people who are misiing native country
```{r}
sum(is.na(unclean_data$native.country))
```

There were 3 countries that shared almost the name ie one was called UnitedStates, Unitedstates and United-States. I replaced UnitedStates and Unitedstates with United-States because there are no countries with such names and I assumed it was a typo that was made.

```{r}
library(plyr)
 revalue( unclean_data$native.country, c("UnitedStates" = "United-States", "Unitedstates" = "United-States" )) ->  unclean_data$native.country
```


```{r}
counts <- table(unclean_data$native.country)
barplot(counts, main = "People in each country ", xlab = " Native countries", cex.names = 0.6, ylab = "frequency", )
```

## 8. Analysing column salaries
I rounded off the salary values to 2 decimal places
```{r}
unclean_data$salaries <- format(round(unclean_data$salaries, 2), nsmall = 2) 
unclean_data$salaries <- as.numeric(unclean_data$salaries)

plot(density(unclean_data$salaries), main = "Salary Distribution",  xlab = "salary", ylab = "Frequency")

```
salary is a ratio.



## 10. Analysing  column Male. 
1 is used to denote males and there 21790 males
```{r}
data.frame(table(unclean_data$male))

```

## 11. Analysing Column  Female

1 is used to analyse females and there 10771 females

```{r}
library(plyr, warn.conflicts = FALSE)

table(unclean_data$female)
```

#Exercise 3

## a)
-I first grouped the data by occupation and education using the group by function

-I then calculated the average salary for each occupation using summarise function

```{r}
library(dplyr, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
library(tidyr, warn.conflicts = FALSE)
data.occupation.education  = group_by(unclean_data, occupation, education)
data.avg.sal               = summarise(data.occupation.education, average_salary=mean(salaries))
data.table <- spread(data.avg.sal, key=education, value=average_salary)
data.table
```

##b
```{r}
data.long <- gather(data.table, education, AvgSalary, "10th":"Some-college")
data.long

```

##c)
```{r}
library(ggplot2, warn.conflicts = FALSE)

ggplot(data.long,  aes(x=education, y=occupation)) +  geom_tile(aes(x=education, y=occupation, fill=AvgSalary))


``` 

##d)
```{r}
#data <- data.frame(lapply(data.long, factor))


data.long$education <- factor(data.long$education, levels = c("Preschool", "1st-4th", "5th-6th", "7th-8th", "9th", "11th", "12th", "HS-grad", "Assoc-voc",  "Assoc-acdm", "levels",
                                                          "Bachelors", "Prof-school", "Masters", "Doctorate" ) , ordered = TRUE)


ggplot(data.long,  aes(x=education, y=occupation)) +  geom_tile(aes(x=education, y=occupation, fill=AvgSalary)) +theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, hjust=1))

```

##e)  List 3 interesting facts that you can read out from this plot.

-Those who ended in preschool can not work in armed forses, protective services, sales, Tech-suppor and transport.
-Those who have masters they atlest everyone has a job.


##f) Create another plot of this dataset that you think conveys interesting information

I first grouped the data by occupation  and the i computed the average salary of each occupation using mean function
```{r}
grouped_data <- group_by(unclean_data, occupation)
summ_data <- summarise(grouped_data, AvgSalary=mean(salaries))
barplot(summ_data$AvgSalary, names.arg=summ_data$occupation, horiz = TRUE, las=1, cex.names = 0.6, xlab = "occupation", ylab = "Average salary", main = "Bar graph showing average salary per department")


```

- on average those in the exec-managerial and prof-speciality department earn more than any other department.
-Those in the riv-house-serv  department earn the least followed by those in the Handlers-cleaners and those in other service department.
-The income difference between departments is not high.


#Exe4
##a
The data has 100000 rows and 15 attributes
```{r}

df <- read.csv("instacart.csv")
str(df)

```

**order_id**
Is of nominal data type. It describes the order of each client who buys a product.

**user_id** is a nominal since its used to identify each person. 

**eval_set**
```{r}
levels(df$eval_set)
levels(as.factor(df$order_dow))
```
its of Nominal attribute. 

**order_number**  its a ratio
```{r}
levels(as.factor(df$order_number))
```

**order_dow** its a ratio because the values are integer values.
```{r}
levels(as.factor(df$order_dow))
```

**order_hour_of_day** its a an ordinal value 
```{r}
levels(as.factor(df$order_hour_of_day))
```

** days_since_prior_order**  its an interval
```{r}
levels(as.factor(df$days_since_prior_order))

```
**product_id** its a nominal value because it uniquely identifies each product

**add_to_cart_order** its a ratio

**reordered** its an ordinal value.

**product_name** its a nominal value because it uniquely identifies the product

**department_id** its 
```{r}
levels(as.factor(df$department))
```

**aisle** its a nominal value that uniquely identifies each aisle
```{r}
levels(as.factor(df$aisle_id))
```
*
**aisle** its a nominal value because each aisle has its own unique name

**department** its a nominal value because each department has its own name
```{r}
levels(df$department)
```

#b
```{r}
library(ggplot2, warn.conflicts = FALSE)
grouped_data <- group_by(df, df$order_hour_of_day)

ggplot(grouped_data, aes(x=grouped_data$order_hour_of_day))+ geom_histogram() + facet_grid(. ~grouped_data$order_dow)

```
##c)  

  -its evident from the histogram that between 10 and 15 hours , there is high turn over of sales for the entire week and sunday having the highes turn over with over 2000 clients.

- its also evident from the histogram that the least sales  in the week are made between 0 and 5 hours.

- its also evident that all the days of the week the sales tend take a similar patter.


##d)
```{r}
grouped_data <- group_by(df, df$order_dow, df$department)

ggplot(grouped_data, aes(x=grouped_data$order_dow))+ geom_histogram() 
```

```{r}
counts <- table( df$department, df$order_dow)
barplot(counts, main="Departmental good bought on different days of the week",
  xlab="Days of the Week", col=c("darkblue","red","yellow","green","orange", "black","DeepPink","Pink", "Lavender","DarkMagenta","Maroon","white","brown","Gainsboro","DarkSlateGray","Chocolate"),
 	legend = rownames(counts))


```

-from the stacked bar graph its obvious that the produce department has the highest sells through the week.

-diary eggs department has the second highest sells during the course of the weeek

-bulk department has the least sells through the week.
